---
title: "Marked Herd Plots"
author: "JM Weissman"
date: "November 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This contains the code to produce plots of Sierra Bighorn population trends. Specifically, these plots are designed to illustrate how the marked populations have changed relative to the herd populations, and how these trends have varied between herds.  

This document is intended to be used to produce population plots and to provide templates for future plots. 

## Requirements:
```{r libraries}
library("tidyverse")
library("ggthemes")
```
## Data Import:
```{r data}
# import csv:
marks_tbl <- read_csv("Overview_Collars.csv")

# replace problem characters in column names:
names(marks_tbl) <- gsub(" ", "_", names(marks_tbl)) 
names(marks_tbl) <- gsub("%", "percent_", names(marks_tbl))

summary(marks_tbl)
```

## Preliminary Data Processing:
```{r processing}
# What % of the population do we want marked?
target_mark_fraction = 0.3

# number of total herds:
n_herds <- dim(unique(marks_tbl['Herd']))[1]

# years over which we have data
years <- seq(min(marks_tbl[['Sheep_Year']]),max(marks_tbl[['Sheep_Year']]))

# Add column for estimated herd total:
marks_tbl <- marks_tbl %>%
  mutate(N_est = round(Marks / percent_Marks))

# mean counts
mean_N <- mean(marks_tbl$N_est, na.rm = T)

# new herds:
new_herds = c("Big Arroyo", "Laurel Creek", "Cathedral", "Bubbs Creek", "Olancha Peak")

# Convert to Long Data

nmarks_tbl <- marks_tbl %>% select(Sheep_Year, Herd, Sex, N_est, Marks, GPS, VHF)
nmarks_tbl <- nmarks_tbl %>% mutate(Mark_Target = as.integer(floor(target_mark_fraction * N_est)) )
nmarks_long_tbl <- nmarks_tbl %>% 
  gather(count_type, n, N_est:Mark_Target) %>%
  mutate(count_type = gsub("N_est", "Est_Total", count_type))
```
## Now the data is in long format, and we can apply filters and produce plots:

```{r plots}
# P1 table: large herds, no vhf, ewes only
p1_tbl <- nmarks_long_tbl %>% 
  filter(Sex == "Female") %>%
  filter(count_type != "VHF") %>%
  group_by(Herd) %>%
  filter(max(n, na.rm = T) > mean_N)

p1 <- ggplot() +
  geom_area(data = p1_tbl %>% filter(count_type != "Mark_Target"),
            aes(x=Sheep_Year, y=n, fill=count_type), position = "identity", alpha = 0.8) + 
  geom_line(data = p1_tbl %>% filter(count_type == "Mark_Target"), 
            aes(x=Sheep_Year, y=n, col="Mark_Target"), size = 1) +
  scale_colour_manual(name = "", values = c("Mark_Target" = "coral3")) +
  scale_fill_manual(name = "Counts", values = 
                      c("Est_Total" = "grey",
                        "Marks" = "cornflowerblue",
                        "GPS" = "blue4")) + 
  scale_x_continuous(expand=c(0,0), limits=c(min(years), max(years))) +
  scale_y_continuous(expand=c(0,0), limits=c(0, max(p1_tbl$n) + 2)) +
  ylab("# of ewes") +
  ggtitle("Sierra Bighorn Marks Relative to Population (large herds only)") + 
  theme(axis.title.x=element_text(size=10, lineheight=.9, face="bold")) + 
  theme(axis.title.y=element_text(size=10, lineheight=.9, face="bold")) + 
  facet_wrap(~ Herd, ncol=2)

p1

# p2 table: new herds, no vhf, ewes only
p2_tbl <- nmarks_long_tbl %>%
  filter(Sex == "Female") %>%
  filter(count_type != "VHF") %>%
  filter(Herd %in% new_herds)

p2 <- ggplot() +
  geom_area(data = p2_tbl %>% filter(count_type != "Mark_Target"),
            aes(x=Sheep_Year, y=n, fill=count_type), position = "identity", alpha = 0.8) + 
  geom_line(data = p2_tbl %>% filter(count_type == "Mark_Target"), 
            aes(x=Sheep_Year, y=n, col="Mark_Target"), size = 1) +
  scale_colour_manual(name = "", values = c("Mark_Target" = "coral3")) +
  scale_fill_manual(name = "Counts", values = 
                      c("Est_Total" = "grey",
                        "Marks" = "cornflowerblue",
                        "GPS" = "blue4")) + 
  scale_x_continuous(expand=c(0,0), limits=c(min(years), max(years))) +
  scale_y_continuous(expand=c(0,0), limits=c(0, max(p2_tbl$n) + 2)) +
  ylab("# of ewes") +
  ggtitle("Sierra Bighorn Marks Relative to Population (new herds only)") + 
  theme(axis.title.x=element_text(size=10, lineheight=.9, face="bold")) + 
  theme(axis.title.y=element_text(size=10, lineheight=.9, face="bold")) + 
  facet_wrap(~ Herd, ncol=2)

p2
```
